---
title: "STAA57 Group Project"
author: "Rebecca Han"
date: "2025-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
df <- read_csv("data/wage-by-education-level.csv")

# Different education levels, other education levels are subsets of the
# main 7 categories

EDUCATION_LEVELS = c(
  "0 - 8  years",
  "Some high school",
  "High school graduate",
  "Some post-secondary",
  "Post-secondary certificate or diploma",
  "Bachelor's degree",
  "Above bachelor's degree"
)
```

https://data.ontario.ca/dataset/wages-by-education-level/resource/7b325fa1-e9d6-4329-a501-08cdc22a79df 

(from crowdmark, requirements)

Description of the variables and the data
- miles

A little background about the data (who collected it in what context)
- bec

What is the over all research question that you are trying to answer in your report
- done

Few tables (appropriately labeled) - bec (do a bunch)
- by age group 
- by province

Few graphs (appropriately labeled)
- by gender (__ graph facet by gender, different colours for [] )

- by year (line graph, different colours for each education level) - miles

```{r}

by_year = df %>% filter(
  `Type of work` == "Both full- and part-time",
  `Education level` %in% EDUCATION_LEVELS,
  Wages == "Median weekly wage rate",
  Geography == "Canada",
  `Age group` == "15 years and over"
)

by_year$`Education level` <- factor(by_year$`Education level`, levels = rev(EDUCATION_LEVELS))

ggplot(by_year, aes(
  x = YEAR,
  y = `Both Sexes`,
  color = `Education level`,
)) + 
  geom_point() + 
  geom_line()

```


/////////////////// 

Use of confidence interval and test of hypothesis

Use of bootstrapping

Use of regression analysis (linear or nonlinear regression [week-9] or logistic regression[week-10] or decision tree/random forest [Week-11]). Interpretation of the regression parameters when applicable.

Use of cross validation

A final summary of your findings

An appendix which shows all your codes

3 marks will be kept aside for the presentation of your report (e.g. proper title, subtitle, graph size, graph placing, line spacing, proper annotation of your code etc.)

TODO: delete above

## TODO: title page 

## Description of the variables and the data

- Wages is aligned with the "both sexes / female / male" columns. 

```{r}
library(tidyverse)
df <- read_csv("data/wage-by-education-level.csv")
colnames(df)[1] = "id" #renames _id to id

#glimpse(df)
df2 <- df %>% filter(Wages %in% c("Average hourly wage rate","Median hourly wage rate"))

#df2 <- df %>% group_by(Wages) %>% summarise(N=n())

df2
```

## A little background about the data (who collected it in what context)

## What is the overall research question that you are trying to answer in your report

Does education level have an impact on someone's wage? 
