---
title: "STAA57 Group Project"
author: "Rebecca Han and Miles Bernstein"
date: "2025-03-19"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
df <- read_csv("data/wage-by-education-level.csv")

# Different education levels, other education levels are subsets of the
# main 7 categories

EDUCATION_LEVELS = c(
  "0 - 8  years",
  "Some high school",
  "High school graduate",
  "Some post-secondary",
  "Post-secondary certificate or diploma",
  "Bachelor's degree",
  "Above bachelor's degree"
)

EDUCATION_LEVELS_AND_ALL = c(
  "Total, all education levels",
  "0 - 8  years",
  "Some high school",
  "High school graduate",
  "Some post-secondary",
  "Post-secondary certificate or diploma",
  "Bachelor's degree",
  "Above bachelor's degree"
)

# Different age groups, other age groups are subsets of the
# main 7 categories

AGE_RANGES = c(
  "15-24 years", 
  "25-54 years", 
  "55 years and over"
)

# https://data.ontario.ca/dataset/wages-by-education-level/resource/7b325fa1-e9d6-4329-a501-08cdc22a79df 
```


## 1 Description of the variables and the data

```{r echo=FALSE}
  colnames(df)
  #df
```


1. **Year**: Year for the wage data
2. **Geography**: Province for the wage data ("Canada" applies for all provinces)
3. **Type of work**: Type of work for the wage data.
  	* Both full- and part-time
	  *	Full-time
	  *	Part-time
4. **Wages**: Type of data collected
	  * Median hourly wage
	  * Average hourly wage
    * Median weekly wage
    * Average weekly wage
    * Total Employees (Number of employees in the specified group, in thousands)
5. **Education level**: Education level for the wage data
    * 0 - 8  years
    * Some high school
    * High school graduate
    * Some post-secondary
    * Post-secondary certificate or diploma
    * Bachelor's degree
    * Above bachelor's degree
6. **Age group**: Age group of for the wage data
7. **Both** Sexes: Wage/Count of Male and Female employees
8. **Male**: Wage/Count of Male employees
9. **Female**: Wage/Count of Female employees


## 2 A little background about the data (who collected it in what context)

The wage data was collected by Ministry of Labour, Immigration, Training and Skills Development
They have been collecting this data since 1997, as part of it's goal to provide neutral 
collective bargaining data, research, and analysis. This is what the Ministry stated about their collection process:

*Beginning January 1997, information is collected on the usual wages or salary of employees at their main job. Respondents are asked to report their wage/salary before taxes and other deductions, and include tips, commissions and bonuses. Weekly and hourly wages/salaries are calculated in conjunction with usual paid work hours per week*


## 3 What is the over all research question that you are trying to answer in your report
Using the wage data from 1997-2019, we are going to investigate the various reasons why wages vary across the population. Our main goal is to discover what factors lead to higher wages and discover trends that may emerge from the data. Our research goals can broken down into three main questions:

1. How much does education level have an impact on someone's wage?
2. What other factors impact wages?
3. How have these factors impacted wages over time?

## 4 Tables 

#### 4.1 Average hourly wage rate and population of each education level
```{r echo=FALSE}
age_edu_count <- df %>% filter(
  `Education level` %in% EDUCATION_LEVELS_AND_ALL,
  `Age group` == "15 years and over",
  YEAR == 2019,
  `Type of work` == "Both full- and part-time",
  Wages %in% c("Average hourly wage rate", "Total employees"),
  Geography == "Canada"
)

age_edu_count2 <- age_edu_count %>% 
  pivot_wider(
    c(`Education level`),
    names_from=`Wages`, 
    values_from=`Both Sexes`
  ) %>%
  mutate(`Total employees` = `Total employees` * 1000)

age_edu_count2 %>%
  rename(
    `Number of Employees` = `Total employees`,
    `Average Hourly Wage` = `Average hourly wage rate`
  ) %>%
  knitr::kable()
```

This table shows the amount of employees, and their average hourly wage for each education level in 2019. There's also an aggregate of all education levels. The table is sorted in order of least, to most formal education. 

#### 4.2 Average hourly wage rate for each age group by education level

```{r echo=FALSE}
# avg hourly wage rate for each age group by education level
age_edu_count <- df %>% filter(
  `Education level` %in% EDUCATION_LEVELS_AND_ALL,
  `Age group` %in% AGE_RANGES,
  YEAR == 2019,
  `Type of work` == "Both full- and part-time",
  Wages == "Average hourly wage rate",
  Geography == "Canada"
)

age_edu_count2 <- age_edu_count %>% 
  pivot_wider(
    c(`Education level`),
    names_from=`Age group`, 
    values_from=`Both Sexes`
  )

age_edu_count2 %>%
  knitr::kable()

```

This table uses the same categorization of education levels, but instead we look at the average hourly wage breakdown by age demographic. 

#### 4.3 Male and Female Wage by education level

```{r echo=FALSE}
age_edu_count <- df %>% filter(
  `Education level` %in% EDUCATION_LEVELS_AND_ALL,
  `Age group` == "15 years and over",
  YEAR == 2019,
  `Type of work` == "Both full- and part-time",
  Wages == "Average hourly wage rate",
  `Geography` == "Canada"
) %>%
  select(`Education level`, Male, Female) %>%
  mutate(`Wage Difference` = Male - Female) %>%
  rename(
    `Average Hourly Male Wage` = Male,
    `Average Hourly Female Wage` = Female
  ) 
  
age_edu_count %>%
  knitr::kable()
```

This table tracks the hourly wage of Males and Females respectively. It also shows how education levels affect the discrepancy. This table also shows the difference between Male and Female hourly wages.

## 5 Graphs 

#### 5.1 Tracking salary by education level over time

```{r echo=FALSE}
by_year = df %>% filter(
  `Type of work` == "Both full- and part-time",
  `Education level` %in% EDUCATION_LEVELS,
  Wages == "Average hourly wage rate",
  Geography == "Canada",
  `Age group` == "15 years and over"
)

by_year$`Education level` <- factor(by_year$`Education level`, levels = rev(EDUCATION_LEVELS))

ggplot(by_year, aes(
  x = YEAR,
  y = `Both Sexes`,
  color = `Education level`,
)) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Average Hourly Wage by Year") +
  xlab("Year") + ylab("Average Hourly Wage")
```
This graph shows the hourly wages of each education level from the years 1997-2019.

## 6 Use of confidence interval and test of hypothesis

Let us now test our hypothesis of the paired t-test $$H_0:$$ There is no difference between Education and Wage vs. $$H_a:$$ There is a difference between Education and Wage

#### 6.1 Hypothesis Test of Average Hourly Wage

```{r echo=FALSE}
df_hourly = df %>% filter(
  `Type of work` == "Both full- and part-time",
  `Education level` %in% EDUCATION_LEVELS,
  Wages == "Average hourly wage rate",
  Geography == "Canada"
) %>%
  mutate(
    Education = case_when(
    `Education level` == "0 - 8 years" ~ 1, 
    `Education level` == "Some high school" ~ 2, 
    `Education level` == "High school graduate" ~ 3, 
    `Education level` == "Some post-secondary" ~ 4, 
    `Education level` == "Post-secondary certificate or diploma" ~ 5, 
    `Education level` == "Bachelor's degree" ~ 6, 
    `Education level` == "Above bachelor's degree" ~ 7, 
    )
  )
  
df_hourly %>%
  with(t.test(x=Education, y=`Both Sexes`, paired = TRUE)) 
```

#### 6.2 Hypothesis Test of Average Weekly Wage

```{r echo=FALSE}
df_weekly = df %>% filter(
  `Type of work` == "Both full- and part-time",
  `Education level` %in% EDUCATION_LEVELS,
  Wages == "Average weekly wage rate",
  Geography == "Canada"
) %>%
  mutate(
    Education = case_when(
    `Education level` == "0 - 8 years" ~ 1, 
    `Education level` == "Some high school" ~ 2, 
    `Education level` == "High school graduate" ~ 3, 
    `Education level` == "Some post-secondary" ~ 4, 
    `Education level` == "Post-secondary certificate or diploma" ~ 5, 
    `Education level` == "Bachelor's degree" ~ 6, 
    `Education level` == "Above bachelor's degree" ~ 7, 
    )
  )
  
df_weekly %>%
  with(t.test(x=Education, y=`Both Sexes`, paired = TRUE)) 
```

## 7 Use of bootstrapping

Since the data from our dataset is aggregated, we should use bootstrapping to ensure that our hypothesis holds even when there are multiple potential samples. 

```{r echo=FALSE}
hourly_df = df %>% filter(
  `Type of work` == "Both full- and part-time",
  `Education level` %in% EDUCATION_LEVELS,
  Wages == "Average hourly wage rate",
  YEAR == "2019"
)

set.seed(100)
#obs.sam=sample(hourly_df,size=10) # Observed sample
boot_function=function(){
  boot_s = sample(hourly_df$`Both Sexes`, size=10, replace=TRUE)
  return(boot_s)
}
boot_X_bar = replicate(100000,boot_function()) 
#95% bootstrap CI
quantile(boot_X_bar, c(0.025,0.975))


```

## 8 Regression analysis 

```{r echo=FALSE}
df_hourly = df %>% filter(
  `Type of work` == "Both full- and part-time",
  `Education level` %in% EDUCATION_LEVELS,
  Wages == "Average hourly wage rate",
  Geography == "Canada",
  YEAR == 2019 #so that we don't have to account for inflation
) %>%
  mutate(
    Education = case_when(
    `Education level` == "0 - 8 years" ~ 1, 
    `Education level` == "Some high school" ~ 2, 
    `Education level` == "High school graduate" ~ 3, 
    `Education level` == "Some post-secondary" ~ 4, 
    `Education level` == "Post-secondary certificate or diploma" ~ 5, 
    `Education level` == "Bachelor's degree" ~ 6, 
    `Education level` == "Above bachelor's degree" ~ 7, 
    )
  )

m <- lm(df_hourly$`Both Sexes` ~ df_hourly$Education)
summary(m)
```

## 9 Use of cross validation

## 10 A final summary of your findings

This table shows a clear trend that more education results in a a higher wage. This comes with an exception, as high school graduates on average have higher wages then those with some post-secondary experience. This table is useful in demonstrating that more education usually leads to higher wages.

Here we see another trend that wages usually increase with age.


## 11 Appendix which shows all your codes


```{r}
library(tidyverse)
df <- read_csv("data/wage-by-education-level.csv")
colnames(df)[1] = "id" #renames _id to id

#glimpse(df)
df2 <- df %>% filter(Wages %in% c("Average hourly wage rate","Median hourly wage rate"))

#df2 <- df %>% group_by(`Age group`) %>% summarise(N=n())

df2
```