---
title: "STAA57 Group Project"
author: "Rebecca Han"
date: "2025-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
df <- read_csv("data/wage-by-education-level.csv")

# Different education levels, other education levels are subsets of the
# main 7 categories

EDUCATION_LEVELS = c(
  "0 - 8  years",
  "Some high school",
  "High school graduate",
  "Some post-secondary",
  "Post-secondary certificate or diploma",
  "Bachelor's degree",
  "Above bachelor's degree"
)

# Different age groups, other age groups are subsets of the
# main 7 categories

AGE_RANGES = c(
  "15-24 years", 
  "25-54 years", 
  "55 years and over"
)
```

https://data.ontario.ca/dataset/wages-by-education-level/resource/7b325fa1-e9d6-4329-a501-08cdc22a79df 

(from crowdmark, requirements)

Description of the variables and the data
- miles

A little background about the data (who collected it in what context)
- bec

What is the over all research question that you are trying to answer in your report
- done

Few tables (appropriately labeled) - bec (do a bunch)
- by age group 
- by province

```{r}
#age_table <- df %>% filter(`Age group` %in% AGE_RANGES) %>% 
#  group_by(Wages, `Age group`) %>%
#  summarize(
#    Wage = mean(`Both Sexes`)
#  ) 
#age_table2 <- age_table %>% pivot_wider(names_from=`Age group`, values_from=Wage)


# avg hourly wage rate for each age group by education level
age_edu_count <- df %>% filter(
  `Education level` %in% EDUCATION_LEVELS,
  `Age group` %in% AGE_RANGES,
  YEAR == 2019,
  `Type of work` == "Both full- and part-time",
  Wages == "Average hourly wage rate",
  Geography == "Canada"
)

age_edu_count2 <- age_edu_count %>% 
  pivot_wider(
    c(`Education level`),
    names_from=`Age group`, 
    values_from=`Both Sexes`
  )
age_edu_count2
```

Few graphs (appropriately labeled)
- by gender (__ graph facet by gender, different colours for [] )

- by year (line graph, different colours for each education level) - miles

```{r}
by_year = df %>% filter(
  `Type of work` == "Both full- and part-time",
  `Education level` %in% EDUCATION_LEVELS,
  Wages == "Median weekly wage rate",
  Geography == "Canada",
  `Age group` == "15 years and over"
)

by_year$`Education level` <- factor(by_year$`Education level`, levels = rev(EDUCATION_LEVELS))

ggplot(by_year, aes(
  x = YEAR,
  y = `Both Sexes`,
  color = `Education level`,
)) + 
  geom_point() + 
  geom_line()

```


/////////////////// 

## Use of confidence interval and test of hypothesis

Let us now test our hypothesis of the paired t-test $$H_0:$$ There is no difference between Education and Wage vs. $$H_a:$$ There is a difference between Education and Wage

- hypothesis test of avg hourly wage

```{r}
df_hourly = df %>% filter(
  `Type of work` == "Both full- and part-time",
  `Education level` %in% EDUCATION_LEVELS,
  Wages == "Average hourly wage rate",
  Geography == "Canada"
) %>%
  mutate(
    Education = case_when(
    `Education level` == "0 - 8 years" ~ 1, 
    `Education level` == "Some high school" ~ 2, 
    `Education level` == "High school graduate" ~ 3, 
    `Education level` == "Some post-secondary" ~ 4, 
    `Education level` == "Post-secondary certificate or diploma" ~ 5, 
    `Education level` == "Bachelor's degree" ~ 6, 
    `Education level` == "Above bachelor's degree" ~ 7, 
    )
  )
  
df_hourly %>%
  with(t.test(x=Education, y=`Both Sexes`, paired = TRUE)) 
```

- hypothesis test of avg weekly wage

```{r}
df_weekly = df %>% filter(
  `Type of work` == "Both full- and part-time",
  `Education level` %in% EDUCATION_LEVELS,
  Wages == "Average weekly wage rate",
  Geography == "Canada"
) %>%
  mutate(
    Education = case_when(
    `Education level` == "0 - 8 years" ~ 1, 
    `Education level` == "Some high school" ~ 2, 
    `Education level` == "High school graduate" ~ 3, 
    `Education level` == "Some post-secondary" ~ 4, 
    `Education level` == "Post-secondary certificate or diploma" ~ 5, 
    `Education level` == "Bachelor's degree" ~ 6, 
    `Education level` == "Above bachelor's degree" ~ 7, 
    )
  )
  
df_weekly %>%
  with(t.test(x=Education, y=`Both Sexes`, paired = TRUE)) 
```

## Use of bootstrapping

Since the data from our dataset is aggregated, we should use bootstrapping to ensure that our hypothesis holds even when there are multiple potential samples. 

```{r}
hourly_df = df %>% filter(
  `Type of work` == "Both full- and part-time",
  `Education level` %in% EDUCATION_LEVELS,
  Wages == "Average hourly wage rate",
  YEAR == "2019"
)

set.seed(100)
#obs.sam=sample(hourly_df,size=10) # Observed sample
boot_function=function(){
  boot_s = sample(hourly_df$`Both Sexes`, size=10, replace=TRUE)
  return(boot_s)
}
boot_X_bar = replicate(100000,boot_function()) 
#95% bootstrap CI
quantile(boot_X_bar, c(0.025,0.975))


```


///////////////////

Use of regression analysis (linear or nonlinear regression [week-9] or logistic regression[week-10] or decision tree/random forest [Week-11]). Interpretation of the regression parameters when applicable.

```{r}
df_hourly = df %>% filter(
  `Type of work` == "Both full- and part-time",
  `Education level` %in% EDUCATION_LEVELS,
  Wages == "Average hourly wage rate",
  Geography == "Canada",
  YEAR == 2019 #so that we don't have to account for inflation
) %>%
  mutate(
    Education = case_when(
    `Education level` == "0 - 8 years" ~ 1, 
    `Education level` == "Some high school" ~ 2, 
    `Education level` == "High school graduate" ~ 3, 
    `Education level` == "Some post-secondary" ~ 4, 
    `Education level` == "Post-secondary certificate or diploma" ~ 5, 
    `Education level` == "Bachelor's degree" ~ 6, 
    `Education level` == "Above bachelor's degree" ~ 7, 
    )
  )

m <- lm(df_hourly$`Both Sexes` ~ df_hourly$Education)
summary(m)
```


Use of cross validation

A final summary of your findings

An appendix which shows all your codes

3 marks will be kept aside for the presentation of your report (e.g. proper title, subtitle, graph size, graph placing, line spacing, proper annotation of your code etc.)

TODO: delete above

## TODO: title page 

## Description of the variables and the data

- Wages is aligned with the "both sexes / female / male" columns. 

```{r}
library(tidyverse)
df <- read_csv("data/wage-by-education-level.csv")
colnames(df)[1] = "id" #renames _id to id

#glimpse(df)
df2 <- df %>% filter(Wages %in% c("Average hourly wage rate","Median hourly wage rate"))

#df2 <- df %>% group_by(`Age group`) %>% summarise(N=n())

df2
```

## A little background about the data (who collected it in what context)

## What is the overall research question that you are trying to answer in your report

Does education level have an impact on someone's wage? 
