---
title: "STAA57 Group Project"
author: "Rebecca Han"
date: "2025-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

https://data.ontario.ca/dataset/wages-by-education-level/resource/7b325fa1-e9d6-4329-a501-08cdc22a79df 

(from crowdmark, requirements)

Description of the variables and the data

A little background about the data (who collected it in what context)

What is the over all research question that you are trying to answer in your report

Few tables (appropriately labeled)

Few graphs (appropriately labeled)

Use of confidence interval and test of hypothesis

Use of bootstrapping

Use of regression analysis (linear or nonlinear regression [week-9] or logistic regression[week-10] or decision tree/random forest [Week-11]). Interpretation of the regression parameters when applicable.

Use of cross validation

A final summary of your findings

An appendix which shows all your codes

3 marks will be kept aside for the presentation of your report (e.g. proper title, subtitle, graph size, graph placing, line spacing, proper annotation of your code etc.)

TODO: delete above

## TODO: title page 

## Description of the variables and the data

- Wages is aligned with the "both sexes / female / male" columns. 

```{r}
library(tidyverse)
df <- read_csv("data/wage-by-education-level.csv")
colnames(df)[1] = "id" #renames _id to id

#glimpse(df)
df2 <- df %>% filter(Wages %in% c("Average hourly wage rate","Median hourly wage rate"))

#df2 <- df %>% group_by(Wages) %>% summarise(N=n())

df2
```

## A little background about the data (who collected it in what context)

## What is the overall research question that you are trying to answer in your report

Does education level have an impact on someone's wage? 
