---
title: "STAA57 Group Project"
author: "Rebecca Han and Miles Bernstein"
date: "2025-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
df <- read_csv("data/wage-by-education-level.csv")

# Different education levels, other education levels are subsets of the
# main 7 categories

EDUCATION_LEVELS = c(
  "0 - 8  years",
  "Some high school",
  "High school graduate",
  "Some post-secondary",
  "Post-secondary certificate or diploma",
  "Bachelor's degree",
  "Above bachelor's degree"
)

# Different age groups, other age groups are subsets of the
# main 7 categories

AGE_RANGES = c(
  "15-24 years", 
  "25-54 years", 
  "55 years and over"
)

# https://data.ontario.ca/dataset/wages-by-education-level/resource/7b325fa1-e9d6-4329-a501-08cdc22a79df 
```


## 1 Description of the variables and the data

```{r echo=FALSE}
  colnames(df)
  #df
```


1. **Year**: Year for the wage data
2. **Geography**: Province for the wage data ("Canada" applies for all provinces)
3. **Type of work**: Type of work for the wage data.
  	* Both full- and part-time
	  *	Full-time
	  *	Part-time
4. **Wages**: Type of data collected
	  * Median hourly wage
	  * Average hourly wage
    * Median weekly wage
    * Average weekly wage
    * Total Employees (Number of employees in the specified group, in thousands)
5. **Education level**: Education level for the wage data
    * 0 - 8  years
    * Some high school
    * High school graduate
    * Some post-secondary
    * Post-secondary certificate or diploma
    * Bachelor's degree
    * Above bachelor's degree
6. **Age group**: Age group of for the wage data
7. **Both** Sexes: Wage/Count of Male and Female employees
8. **Male**: Wage/Count of Male employees
9. **Female**: Wage/Count of Female employees


## 2 A little background about the data (who collected it in what context)

The wage data was collected by Ministry of Labour, Immigration, Training and Skills Development
They have been collecting this data since 1997, as part of it's goal to provide neutral 
collective bargaining data, research, and analysis. This is what the Ministry stated about their collection process:

*Beginning January 1997, information is collected on the usual wages or salary of employees at their main job. Respondents are asked to report their wage/salary before taxes and other deductions, and include tips, commissions and bonuses. Weekly and hourly wages/salaries are calculated in conjunction with usual paid work hours per week*


## 3 What is the over all research question that you are trying to answer in your report
Does education level have an impact on someone's wage? 

## 4 Tables 

#### 4.1 Average hourly wage rate for each age group by education level

```{r echo=FALSE}
#age_table <- df %>% filter(`Age group` %in% AGE_RANGES) %>% 
#  group_by(Wages, `Age group`) %>%
#  summarize(
#    Wage = mean(`Both Sexes`)
#  ) 
#age_table2 <- age_table %>% pivot_wider(names_from=`Age group`, values_from=Wage)


# avg hourly wage rate for each age group by education level
age_edu_count <- df %>% filter(
  `Education level` %in% EDUCATION_LEVELS,
  `Age group` %in% AGE_RANGES,
  YEAR == 2019,
  `Type of work` == "Both full- and part-time",
  Wages == "Average hourly wage rate",
  Geography == "Canada"
)

age_edu_count2 <- age_edu_count %>% 
  pivot_wider(
    c(`Education level`),
    names_from=`Age group`, 
    values_from=`Both Sexes`
  )

age_edu_count2 %>%
  knitr::kable()

```

#### 4.2 by province (unsure if needed) 
#### 4.3 by gender (unsure if needed)

## 5 Graphs 

#### 5.1 Tracking salary by education level over time

```{r echo=FALSE}
by_year = df %>% filter(
  `Type of work` == "Both full- and part-time",
  `Education level` %in% EDUCATION_LEVELS,
  Wages == "Median weekly wage rate",
  Geography == "Canada",
  `Age group` == "15 years and over"
)

by_year$`Education level` <- factor(by_year$`Education level`, levels = rev(EDUCATION_LEVELS))

ggplot(by_year, aes(
  x = YEAR,
  y = `Both Sexes`,
  color = `Education level`,
)) + 
  geom_point() + 
  geom_line()

```

#### 5.2 Total salary of different education group (unsure if needed)



## 6 Use of confidence interval and test of hypothesis

Let us now test our hypothesis of the paired t-test $$H_0:$$ There is no difference between Education and Wage vs. $$H_a:$$ There is a difference between Education and Wage

- hypothesis test of avg hourly wage

```{r}
df_hourly = df %>% filter(
  `Type of work` == "Both full- and part-time",
  `Education level` %in% EDUCATION_LEVELS,
  Wages == "Average hourly wage rate",
  Geography == "Canada"
) %>%
  mutate(
    Education = case_when(
    `Education level` == "0 - 8 years" ~ 1, 
    `Education level` == "Some high school" ~ 2, 
    `Education level` == "High school graduate" ~ 3, 
    `Education level` == "Some post-secondary" ~ 4, 
    `Education level` == "Post-secondary certificate or diploma" ~ 5, 
    `Education level` == "Bachelor's degree" ~ 6, 
    `Education level` == "Above bachelor's degree" ~ 7, 
    )
  )
  
df_hourly %>%
  with(t.test(x=Education, y=`Both Sexes`, paired = TRUE)) 
```

- hypothesis test of avg weekly wage

```{r}
df_weekly = df %>% filter(
  `Type of work` == "Both full- and part-time",
  `Education level` %in% EDUCATION_LEVELS,
  Wages == "Average weekly wage rate",
  Geography == "Canada"
) %>%
  mutate(
    Education = case_when(
    `Education level` == "0 - 8 years" ~ 1, 
    `Education level` == "Some high school" ~ 2, 
    `Education level` == "High school graduate" ~ 3, 
    `Education level` == "Some post-secondary" ~ 4, 
    `Education level` == "Post-secondary certificate or diploma" ~ 5, 
    `Education level` == "Bachelor's degree" ~ 6, 
    `Education level` == "Above bachelor's degree" ~ 7, 
    )
  )
  
df_weekly %>%
  with(t.test(x=Education, y=`Both Sexes`, paired = TRUE)) 
```

## 7 Use of bootstrapping

Since the data from our dataset is aggregated, we should use bootstrapping to ensure that our hypothesis holds even when there are multiple potential samples. 

```{r}
hourly_df = df %>% filter(
  `Type of work` == "Both full- and part-time",
  `Education level` %in% EDUCATION_LEVELS,
  Wages == "Average hourly wage rate",
  YEAR == "2019"
)

set.seed(100)
#obs.sam=sample(hourly_df,size=10) # Observed sample
boot_function=function(){
  boot_s = sample(hourly_df$`Both Sexes`, size=10, replace=TRUE)
  return(boot_s)
}
boot_X_bar = replicate(100000,boot_function()) 
#95% bootstrap CI
quantile(boot_X_bar, c(0.025,0.975))


```


///////////////////

Use of regression analysis (linear or nonlinear regression [week-9] or logistic regression[week-10] or decision tree/random forest [Week-11]). Interpretation of the regression parameters when applicable.

```{r}
df_hourly = df %>% filter(
  `Type of work` == "Both full- and part-time",
  `Education level` %in% EDUCATION_LEVELS,
  Wages == "Average hourly wage rate",
  Geography == "Canada",
  YEAR == 2019 #so that we don't have to account for inflation
) %>%
  mutate(
    Education = case_when(
    `Education level` == "0 - 8 years" ~ 1, 
    `Education level` == "Some high school" ~ 2, 
    `Education level` == "High school graduate" ~ 3, 
    `Education level` == "Some post-secondary" ~ 4, 
    `Education level` == "Post-secondary certificate or diploma" ~ 5, 
    `Education level` == "Bachelor's degree" ~ 6, 
    `Education level` == "Above bachelor's degree" ~ 7, 
    )
  )

m <- lm(df_hourly$`Both Sexes` ~ df_hourly$Education)
summary(m)
```


## 8 Use of cross validation

## 9 A final summary of your findings

## Appendix which shows all your codes


```{r}
library(tidyverse)
df <- read_csv("data/wage-by-education-level.csv")
colnames(df)[1] = "id" #renames _id to id

#glimpse(df)
df2 <- df %>% filter(Wages %in% c("Average hourly wage rate","Median hourly wage rate"))

#df2 <- df %>% group_by(`Age group`) %>% summarise(N=n())

df2
```